GVN-Hoist: Hoisting Computations from Branches

### 2016-08-11 hoist.tex
memref
cost models

\section{Cost models}

\subsubsection{Reduce register pressure}
\label{hoist:reg-pressure}
Following example explains how code hoisting can actually reduce the register
pressure.  Consider the following example where the labels prefixed with 'P'
represent the position of instruction in a basic block (names prefixed with 'B').

@@

\subsubsection{In the presence of calls}
\label{cost:across-calls}
Hoisting scalars across calls is tricky because it can increase the number of
spills. During the frame lowering of calls, the argument registers, in general,
the caller saved registers are saved because they might be modified by the
callee and after the call they are restored \cite{frame-lowering}. So before the
call, the register pressure is high because the number of available registers
are reduced by the number of caller saved registers. In that situation if a
computation is hoisted across the call, that would increase the total number of
registers required by one, thus contributing to the register pressure.

@@

Hoisting memory references also require precise analysis of all the memory
addresses accessed by the call. In our implementation, since we only analyze the
callee declaration the analysis is very conservative. In the presence of pure
calls, loads can be hoisted but stores can't. Also, if the call throws exceptions,
or it it may not return, memory references cannot be hoisted.

\subsubsection{Hoisting too far away}
When we load a value from memory, we require an extra register to keep the value
around. Hoisting a load means we need to keep the value alive for a longer
period. That could result in spills later in the register allocation.

Also, if there are several instructions in between the hoisting point and the
instruction to be hoisted, the instruction to be hoisted crosses several
instructions while hoisting, it means we are adding one register to all the
live-ranges spanning the instructions. That could result in spills. In the
current implementation we choose to hoist if the number of instructions crossed
is below a threshold. Ideally, it should be okay to hoist all the instructions
and a later live-range-splitting pass should make the right decision of
rematerializing the instruction should it be beneficial to do so. But the
current live-range splitting pass of llvm is not making the optimial decision
and we have found spills if the threshold is exceeded. The threshold was
computed as a result of tuning the llvm testsuite \cite{llvm-nightly} and spec
benchmark \cite{Henning2000}.

\begin{enumerate}
\item When the hoist point is too far.
\item When the number of instructions we may cross is very high.
\end{enumerate}

>>>>

\section{Cost models}
Similar to any compiler optimization pass, there are several cost functions that
are deployed to tune for optimal combination of performance and code-size.
Since this is mostly a code-size optimization pass, the goal is to not regress
in performance across popular benchmarks at the same time reduce code size as
much as possible. Following are the cost models which are implemented:

\subsection{Reduce register pressure}
\label{hoist:reg-pressure}
Following example explains how code hoisting can actually reduce the register
pressure.  Consider the following example where the labels prefixed with 'P'
represent the position of instruction in a basic block (names prefixed with 'B').

@@

\subsection{In the presence of calls}
\label{cost:across-calls}
Hoisting scalars across calls is tricky because it can increase the number of
spills. During the frame lowering of calls, the argument registers, in general,
the caller saved registers are saved because they might be modified by the
callee and after the call they are restored \cite{frame-lowering}. So before the
call, the register pressure is high because the number of available registers
are reduced by the number of caller saved registers. In that situation if a
computation is hoisted across the call, that would increase the total number of
registers required by one, thus contributing to the register pressure. However,
in the special case discussed in Section-\ref{hoist:reg-pressure}, it will be
okay to hoist because the register pressure would not decrease.

Hoisting loads/stores across calls also require precise analysis of all the
memory addresses accessed by the call. Our implementation being an
intraprocedural pass, the analysis is very conservative. In the presence of pure
calls, loads can be hoisted but stores can't. Also, if the call throws
exceptions, or it it may not return, memory references cannot be hoisted.

\subsection{Hoisting too far away}
If there are several instructions in between the hoisting point and the
instruction to be hoisted, the instruction to be hoisted crosses several
instructions while hoisting, that means we are adding one register to all the
live-ranges spanning the instructions. That could result in spills. In the
current implementation we choose to hoist if the number of instructions crossed
is below a threshold. Ideally, it should be okay to hoist all the instructions
and a later a live-range-splitting \cite{cooper1998live} pass should make the
right decision of rematerializing the instruction should it be beneficial to do
so. But the current live-range splitting pass of llvm is not making the optimial
decision and we have found spills if the threshold is exceeded. The threshold
was computed as a result of tuning the llvm testsuite \cite{llvm-nightly} and
spec benchmark \cite{Henning2000}.

Also, hoisting a load increases the register pressure by one across all the
instructions which the load would cross. That could result in spills later in
the register allocation.

@@
